---
title: "R Notebook"
output: html_notebook
---

define used libraries

```{r}
libraries_used <- c("lazyeval", "readr","plyr" ,"dplyr", "readxl", "ggplot2", 
    "funModeling", "scales", "tidyverse", "corrplot", "GGally", "caret",
    "rpart", "randomForest", "pROC", "gbm", "choroplethr", "choroplethrMaps",
    "microbenchmark", "doParallel", "e1071")
```

check missing libraries and install them

```{r}
libraries_used[!(libraries_used %in% installed.packages()[,"Package"])]
if(length(libraries_missing)) install.packages(libraries_missing)
```

use dplyr
```{r}
library(dplyr)
```

load dataset
```{r}
df <- read.csv("regression_train_loan.csv")
```

get meta information and output it
```{r}
meta_df <- funModeling::df_status( df, print_results = FALSE )

knitr::kable(meta_df)
```

get amount and percentage of unique and output the table
```{r}
meta_df_p <- meta_df %>% mutate(uniq_rat = unique / nrow(df))
meta_df_p %>% select(variable, unique, uniq_rat) %>% mutate(unique = unique, uniq_rat = scales::percent(uniq_rat)) %>% knitr::kable()
```

convert date attributes
```{r}
chr_to_date_vars <-  c("issue_d", "last_pymnt_d", "last_credit_pull_d", "next_pymnt_d", "earliest_cr_line", "next_pymnt_d")
convert_date <- function(x){
  as.Date(paste0("01-", x), format = "%d-%b-%Y")
}

df <- df %>% mutate_at(.funs = funs(convert_date), .vars = chr_to_date_vars)
```

output a table with amount of unique, N/A, and zero values
```{r}
meta_df %>% select(variable, p_zeros, p_na, unique) %>% knitr::kable()
```

recreate meta table
```{r}
meta_df <- funModeling::df_status( df, print_results = FALSE )
```

set vars from N/A to zero
```{r}
na_to_zero_vars <- c("mths_since_last_delinq", "mths_since_last_record", "mths_since_last_major_derog")

df <- df %>% mutate_at(.vars = na_to_zero_vars, .funs = funs(replace(., is.na(.), 0)))
```

remove columns deemed unfit for modeling
```{r}
vars_to_remove <- c("annual_inc_joint", "dti_joint", "policy_code", "id", "member_id",
                    "emp_title", "url", "desc", "title", "open_acc_6m", "open_il_6m", 
                    "open_il_12m", "open_il_24m", "mths_since_rcnt_il", "total_bal_il", 
                    "il_util", "open_rv_12m", "open_rv_24m", "max_bal_bc", "all_util",
                    "total_rev_hi_lim", "inq_fi", "total_cu_tl", "inq_last_12m",
                    "verification_status_joint", "next_pymnt_d", "sub_grade", "X")

df_clean <- df %>% select(-one_of(vars_to_remove))
```

set the seed for reproductionability
```{r}
set.seed(1)
```

set the training and test data (80% train, 20% test)
```{r}
train_index <- caret::createDataPartition(y = df_clean$int_rate, times = 1, p = .8, list = FALSE)

train <- df_clean[train_index, ]
test <- df_clean[-train_index, ]
```


Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
